---
title       : text analyst
subtitle    : about shiny application for text analysis
author      : R.G.
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [bootstrap, shiny, interactive]   # {mathjax, quiz, bootstrap}
mode        : selfcontained     # {standalone, draft}
knit        : slidify::knit2slides
---

## Application and its structure

* text analyst is a shiny application
* find it on https://rgrbic.shinyapps.io/appWords
* user enters text to be analyzed, app shows word frequency 
* user interface is <code>sidebarLayout</code> which consists of <code>sidebarPanel</code> and <code>mainPanel</code>
* <code>sidebarPanel</code> contains short application description with the radio buttons where user selects desired plot
* <code>mainPanel</code> is divided into two rows:
    + first row is input text area where user enters text to be analyzed
    + second row is output, i.e. plot of word frequency
* two plots are available:
    + word cloud plot that uses R <code>wordcloud</code> library
    + barplot of 10 most common words in the text, made with R <code>ggplot2</code> library

--- .class #id 

## Under the hood

* the code that processes the text from input text area is in <code>server.R</code>
* it first transforms text to lowercase and then removes all special charachters
* after that text is splitted into separate words and number of occurences of each word is calculated

```{r uhood, echo = TRUE, eval = FALSE}
getWords <- function(text_object){
    
    text <- tolower(text_object)
    textF <- gsub("[[:punct:]]", " ", text)
    textF <- gsub("\r\n|\r|\n", " ", textF)
    textSplitted <- strsplit(textF," +")
    textWords <- unlist(textSplitted)    
    words <- plyr::count(textWords)    
    wordsSorted <- words[order(-words$freq),]
    
    wordsSorted
}
```

--- .class #id


## Word Cloud

* for plotting word cloud R library <code>wordcloud</code> is used:

```{r wcloud, echo = FALSE, eval = TRUE, fig.align='center', message=FALSE}
library(wordcloud)

text_object = "Two households, both alike in dignity,
In fair Verona, where we lay our scene,
From ancient grudge break to new mutiny,
Where civil blood makes civil hands unclean.
From forth the fatal loins of these two foes
A pair of star-cross'd lovers take their life;
Whose misadventured piteous overthrows
Do with their death bury their parents' strife.
The fearful passage of their death-mark'd love,
And the continuance of their parents' rage,
Which, but their children's end, nought could remove,
Is now the two hours' traffic of our stage;
The which if you with patient ears attend,
What here shall miss, our toil shall strive to mend."

text <- tolower(text_object)
textF <- gsub("[[:punct:]]", " ", text)
textF <- gsub("\r\n|\r|\n", " ", textF)
textSplitted <- strsplit(textF," +")
textWords <- unlist(textSplitted)    
words <- plyr::count(textWords)    
wordsSorted <- words[order(-words$freq),]
wordcloud(wordsSorted$x,wordsSorted$freq,scale=c(6,.5),min.freq=2,max.words=Inf)

```

--- .class #id

## Barplot of word occurrences

* for plotting barplot R library <code>ggplot2</code> is used

```{r barp, echo = FALSE, eval = TRUE, fig.align='center', warnings = FALSE, message=FALSE}
library(ggplot2)

text_object = "Two households, both alike in dignity,
In fair Verona, where we lay our scene,
From ancient grudge break to new mutiny,
Where civil blood makes civil hands unclean.
From forth the fatal loins of these two foes
A pair of star-cross'd lovers take their life;
Whose misadventured piteous overthrows
Do with their death bury their parents' strife.
The fearful passage of their death-mark'd love,
And the continuance of their parents' rage,
Which, but their children's end, nought could remove,
Is now the two hours' traffic of our stage;
The which if you with patient ears attend,
What here shall miss, our toil shall strive to mend."

text <- tolower(text_object)
textF <- gsub("[[:punct:]]", " ", text)
textF <- gsub("\r\n|\r|\n", " ", textF)
textSplitted <- strsplit(textF," +")
textWords <- unlist(textSplitted)    
words <- plyr::count(textWords)    
wordsSorted <- words[order(-words$freq),]

ggplot(data = wordsSorted[1:10,], aes(x=reorder(x, freq), y = freq)) + 
    geom_bar(stat = "identity", fill = "steelblue") + 
    coord_flip() + 
    labs(y = "") + 
    theme(axis.text.y = element_text(colour="grey20", size=14),
          axis.text.x = element_text(colour="grey20", size=14)) +
    labs(x = "")
```

